---
title: "Sheet 01 — data cleaning, groups, boxplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 01 — data cleaning, groups, boxplot

**Data:** `data/Polisy.csv`, `data/Korozja.csv`. **Time:** ~35 min.

Set working directory to this folder (RStudio: Session → Set Working Directory → To Source
File Location). Solutions: `rozwiazania.Rmd`.

> These are Polish-locale files (separator `;`, decimal comma) — pay attention to which
> read function you use.

**Ex. 1.** Load `data/Polisy.csv`. How many observations (rows) are there? What will happen
if you use `read.csv()` instead of `read.csv2()`?

```{r}
```

**Ex. 2.** What is the mean driver age (`Wiek_kierowcy`) separately for women and men?

```{r}
```

**Ex. 3.** What percentage of policies ended in a claim (`Szkoda == "Tak"`)?

```{r}
```

**Ex. 4.** Do drivers with a claim have a longer or shorter mean usage time
(`Czas_eksploatacji`) than those without a claim? Report both means.

```{r}
```

**Ex. 5.** Load `data/Korozja.csv`. Draw a boxplot of damage depth by coating type.
What is the **median** depth for coating A, and for B? Is the thick line in the box the
median or the mean?

```{r}
```

**Ex. 6.** Calculate the mean driver age broken down by `Plec` **and** `Szkoda`
simultaneously (4 groups). What surprising thing do you see for men?

```{r}
```

**Ex. 7.** What is the correlation between `Wiek_kierowcy` and `Czas_eksploatacji`?

```{r}
```

**Ex. 8.** Among policies with a claim (`Szkoda == "Tak"`), what percentage are women?

```{r}
```

**Ex. 9.** For `Korozja.csv`: which coating has greater **spread** of damage depth?
Compare the standard deviation and IQR for A and B.

```{r}
```

**Ex. 10.** Draw a boxplot of depth by coating in **ggplot2**, with coating on the x-axis
and `fill` by coating.

```{r}
```

**Ex. 11.** (closed) Which function in `tidymodels` splits data into training and test sets?

- `split_test_train()`
- `train_test_split()`
- `initial_split()`
- `data_split()`

**Ex. 12.** (closed) What arguments should you use in `facet_grid()` to split a plot into
a grid by two variables `var1` and `var2` with "free" scales?

- `facet_grid(var1 + var2, scales="free")`
- `facet_grid(var1 ~ var2, scales="free")`
- `facet_grid(var1 | var2, scales="free")`

**Ex. 13.** (closed) What is the key property of an atomic vector in R?

- elements of different classes
- elements of the same class
- length < 32768
- no names
