---
title: "Sheet 01 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 01 — solutions

**Ex. 1.**

```{r}
pol <- read.csv2("data/Polisy.csv") # ; separator + , decimal
nrow(pol) # 348
```

**Ans.: 348 observations.** With `read.csv()`, R would treat `,` (from numbers like `48,6`) as
the separator → error "more columns than column names" / numbers read as text.

**Ex. 2.**

```{r}
library(dplyr)
pol %>%
  group_by(Plec) %>%
  summarise(sredni_wiek = mean(Wiek_kierowcy))
# base R: tapply(pol$Wiek_kierowcy, pol$Plec, mean)
```

**Ans.: Female ≈ 39.7; Male ≈ 46.6.**

**Ex. 3.**

```{r}
mean(pol$Szkoda == "Tak") * 100 # 20.40 %
sum(pol$Szkoda == "Tak") # 71
```

**Ans.: ≈ 20.4% (71 of 348).**

**Ex. 4.**

```{r}
pol %>%
  group_by(Szkoda) %>%
  summarise(sredni_czas = mean(Czas_eksploatacji))
```

**Ans.: with claim 8.03 vs without claim 4.74 → with claim LONGER usage time.**

**Ex. 5.**

```{r}
kor <- read.csv2("data/Korozja.csv")
names(kor) # "Rodzaj.powłoki" "Głębokość.uszkodzenia"
boxplot(Głębokość.uszkodzenia ~ Rodzaj.powłoki, data = kor)
tapply(kor$Głębokość.uszkodzenia, kor$Rodzaj.powłoki, median)
```

**Ans.: median A = 55, B = 44. The line in the box is the MEDIAN.** (means: A 54.2; B 46.2)

**Ex. 6.**

```{r}
pol %>%
  group_by(Plec, Szkoda) %>%
  summarise(n = n(), wiek = mean(Wiek_kierowcy), .groups = "drop")
```

**Ans.: Female/No 39.8; Female/Yes 38.8; Male/No 49.9; Male/Yes 35.4.**
Surprising: men **with a claim** are much YOUNGER (35.4) than those without (49.9).

**Ex. 7.**

```{r}
cor(pol$Wiek_kierowcy, pol$Czas_eksploatacji) # -0.244
```

**Ans.: ≈ −0.24 (weak negative).**

**Ex. 8.**

```{r}
mean(pol$Plec[pol$Szkoda == "Tak"] == "Kobieta") * 100 # 15.49
```

**Ans.: ≈ 15.5%.**

**Ex. 9.**

```{r}
tapply(kor$Głębokość.uszkodzenia, kor$Rodzaj.powłoki, sd) # A 11.58 ; B 10.77
tapply(kor$Głębokość.uszkodzenia, kor$Rodzaj.powłoki, IQR) # A 12.5  ; B 11.5
```

**Ans.: coating A has greater spread.**

**Ex. 10.**

```{r}
library(ggplot2)
ggplot(kor, aes(x = Rodzaj.powłoki, y = Głębokość.uszkodzenia, fill = Rodzaj.powłoki)) +
  geom_boxplot()
```

**Ex. 11.** → **`initial_split()`** (then `training()` / `testing()`).

**Ex. 12.** → **`facet_grid(var1 ~ var2, scales = "free")`** (tilde separates the variables).

**Ex. 13.** → **elements of the same class.**
