---
title: "Sheet 10 — TWO datasets: hospitals (treatment outcomes + hospital data)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 10 — TWO datasets: hospitals (treatment outcomes + hospital data)

**Data:** `data/outcome-of-care-measures.csv` (mortality), `data/hospital-data.csv`
(type/ownership/location). **Time:** ~40 min. **Level: medium→hard.**

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> Join on `Provider.Number`. Mortality columns are text with `"Not Available"` →
> `as.numeric()` creates `NA`. Conditions: col. 11 = heart attack (HA), 17 = heart failure
> (HF), 23 = pneumonia (PN).

**Ex. 1.** Join both datasets (inner join) on `Provider.Number`. How many records match?

```{r}
```

**Ex. 2.** How many hospitals in `hospital-data.csv` have NO match in `outcome` (anti-join)?
And how many `outcome` records have no match in `hospital-data`?

```{r}
```

**Ex. 3.** After joining, compute the mean **pneumonia** mortality (`pn`) by `Hospital.Type`.
Which type has the lowest mean?

```{r}
```

**Ex. 4.** Which `Hospital.Ownership` has the lowest mean **heart failure** (`hf`) mortality?

```{r}
```

**Ex. 5.** For each state, find the hospital with the lowest **heart attack** (`ha`)
mortality. In how many states is that best hospital of type `"Acute Care Hospitals"`?

```{r}
```

**Ex. 6.** (closed) `anti_join(a, b, by="k")` returns:

- rows of `a` that HAVE a match in `b`
- rows of `a` that DO NOT have a match in `b`
- shared rows

**Ex. 7.** (closed) To keep, for each group, the row with the smallest value of `x`:

- `filter(min(x))`
- `group_by() %>% slice_min(x, n=1)`
- `summarise(min(x))`

**Ex. 8.** (closed) `left_join(a, b)` as opposed to `inner_join(a, b)`:

- keeps only shared keys
- keeps all rows of `a` (missing values = NA)
- keeps all rows of `b`
