---
title: "Sheet 07 — multiple files, loops, custom functions (specdata)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 07 — multiple files, loops, custom functions (specdata)

**Data:** `data/specdata/` — files `001.csv` … `332.csv` (each: `Date`, `sulfate`,
`nitrate`, `ID`). **Time:** ~45 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> This type of task comes from the "air pollution" quiz: read many files and operate on
> the combined data.

**Ex. 1.** Write a function `pollutantmean(directory, pollutant, id = 1:332)` that
reads the files with numbers `id`, combines them, and returns the mean of the specified
pollutant (ignoring `NA`). What is `pollutantmean("data/specdata", "sulfate", 1:10)`?
(round to 3 decimal places)

```{r}
```

**Ex. 2.** What is `pollutantmean("data/specdata", "nitrate", 70:72)`? (3 decimal places)

```{r}
```

**Ex. 3.** Write a function `complete(directory, id)` that returns a data frame `id, nobs`
(number of complete rows). How many complete observations does file 54 have?

```{r}
```

**Ex. 4.** Using `complete`, how many complete observations do files 1–10 have in total?

```{r}
```

**Ex. 5.** Write a function `corr(directory, threshold = 0)` that, for each monitor with
more than `threshold` **complete** observations, computes the `sulfate` vs `nitrate`
correlation and returns a vector of those correlations. What is `length(corr("data/specdata", 150))`?

```{r}
```

**Ex. 6.** What are `length(corr("data/specdata", 400))` and `length(corr("data/specdata", 1000))`?

```{r}
```

**Ex. 7.** What is the **mean** correlation for threshold 150?

```{r}
```

**Ex. 8.** (closed) How do you get all file paths in a directory?

- `read.csv("specdata")`
- `list.files("specdata", full.names = TRUE)`
- `dir.exists("specdata")`

**Ex. 9.** (closed) How do you combine a list of data frames `lista` into one data frame?

- `do.call(rbind, lista)`
- `rbind(lista)`
- `merge(lista)`

**Ex. 10.** (closed) `complete.cases(df)` returns:

- the number of columns
- a logical vector TRUE for rows without `NA`
- a data frame without `NA`

**Ex. 11.** (closed) To compare the speed of two function implementations, you use:

- `profvis()`
- `microbenchmark()`
- `traceback()`

**Ex. 12.** (closed) In the `corr` function, `complete.cases(d)` is used to:

- remove columns with NA
- select rows without `NA` (for computing the correlation)
- count NA values
