---
title: "Sheet 13 — PM2.5 emissions: two RDS files, join, aggregation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Data:** `data/summarySCC_PM25.rds` (6.5 million rows) and `data/Source_Classification_Code.rds`
(11,717 rows). **Time:** ~45 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> `summarySCC_PM25.rds`: PM2.5 particulate emissions in the USA for years 1999, 2002, 2005, 2008.
> Columns: `fips` (county code), `SCC` (source code), `Pollutant`, `Emissions`, `type`, `year`.
> `Source_Classification_Code.rds`: source dictionary — `SCC`, `EI.Sector`, `SCC.Level.One`…`SCC.Level.Four`.
> Join key: column `SCC`.
> Loading: `readRDS("file.rds")` — the object is assigned to a variable (unlike `load()`).

**Ex. 1.** Load both files. Check `dim()` and `names()` of each. What years does
`summarySCC_PM25.rds` contain? What emission types (`type`)?

```{r}
```

**Ex. 2.** Compute total emissions (`sum(Emissions)`) for each year. Is the trend
increasing or decreasing?

```{r}
```

**Ex. 3.** How much did total emissions in 1999 and in 2008? By what percentage did they fall?

```{r}
```

**Ex. 4.** For Baltimore (`fips == "24510"`), compute total emissions by year. Does
Baltimore follow the national trend?

```{r}
```

**Ex. 5.** Which `type` (POINT / NONPOINT / ON-ROAD / NON-ROAD) emits the most in
Baltimore in 1999? And in 2008? Did the proportions change?

```{r}
```

**Ex. 6.** Coal combustion: select from the SCC dictionary rows where `SCC.Level.One`
contains "Combustion" **and** `SCC.Level.Four` contains "Coal". Join with NEI and compute
emissions per year. How many SCC codes match this criterion?

```{r}
```

**Ex. 7.** Motor vehicles in Baltimore: filter NEI for `fips == "24510"` and
`type == "ON-ROAD"`. Compute emissions per year. Are road traffic emissions rising or falling?

```{r}
```

**Ex. 8.** Top 5 sectors (`EI.Sector`) by total emissions (all years, entire USA).
Requires joining NEI with SCC.

```{r}
```

**Ex. 9.** Compare the ON-ROAD emissions trend in Baltimore (`fips == "24510"`) vs
Los Angeles (`fips == "06037"`) per year. Which city has higher emissions in 2008?

```{r}
```

**Ex. 10.** Which `SCC` codes appear in NEI but have **no** match in the SCC table?
Use `anti_join`. How many such observations are there?

```{r}
```

**Ex. 11.** (closed) How do you load a `.rds` file?

- `load("file.rds")`
- `x <- readRDS("file.rds")`
- `x <- read.rds("file.rds")`
- `source("file.rds")`

**Ex. 12.** (closed) `anti_join(A, B, by="SCC")` returns:

- rows of A that have a match in B
- rows of B that have no match in A
- rows of A that **have no** match in B
- all rows from A and B

**Ex. 13.** (closed) After `load("file.rda")` the object:

- is returned as a value — you must assign it: `x <- load(...)`
- appears in the environment under its own name automatically
- is loaded into a list named `file`

**Ex. 14.** For each `EI.Sector`, compute emissions in 1999 and 2008 and their ratio
(2008 / 1999). Which sectors reduced emissions the most (lowest ratio)?

```{r}
```
