---
title: "Sheet 06 — dates, sorting, extraction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 06 — dates, sorting, extraction

**Data:** `data/gaData.csv` (columns: `date` as text "YYYY-MM-DD", `visitsTumblr`). **Time:** ~35 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

**Ex. 1.** Load `data/gaData.csv`. Convert `date` to class `Date`. What is the date range
(from–to)? What is the total number of visits?

```{r}
```

**Ex. 2.** On how many days was the number of visits greater than 0?

```{r}
```

**Ex. 3.** How many visits were recorded in total in **2012**? (Extract the year from the date.)

```{r}
```

**Ex. 4.** Which **day of the week** has the highest total visits?

```{r}
```

**Ex. 5.** Which **month** (in YYYY-MM format) had the highest total visits?

```{r}
```

**Ex. 6.** Which single **day** had the most visits and how many?

```{r}
```

**Ex. 7.** How many days belong to 2011, and how many to 2012?

```{r}
```

**Ex. 8.** (closed) How do you extract the year from a `Date` object named `d`?

- `year(d)` from base R
- `format(d, "%Y")`
- `substr(d, 1, 2)`
- `d$year`

**Ex. 9.** (closed) What does `weekdays(as.Date("2012-01-02"))` do?

- returns the day number
- returns the name of the day of the week
- returns the month

**Ex. 10.** (closed) Correct order for saving a plot to PDF:

- `pdf("f.pdf")` → code → `dev.off()`
- code → `pdf("f.pdf")` → `dev.off()`
- `dev.on("f.pdf")` → code → `dev.off()`

**Ex. 11.** (closed) The function `traceback()` is used to:

- set a breakpoint
- show the call stack after an error
- measure time

**Ex. 12.** (closed) To extract just the year (4 characters) from the string "2012-01-02":

- `substr(x, 1, 4)`
- `str_extract(x, "^[0-9]{4}")`
- both work
