---
title: "Sheet 06 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 06 — solutions

```{r}
ga <- read.csv("data/gaData.csv")
ga$date <- as.Date(ga$date) # ISO format -> parsed automatically
```

**Ex. 1.**

```{r}
range(ga$date) # "2011-01-01" "2012-08-22"
sum(ga$visitsTumblr) # 124156
```

**Ans.: 2011-01-01 → 2012-08-22; total = 124156.**

**Ex. 2.**

```{r}
sum(ga$visitsTumblr > 0) # 350
```

**Ans.: 350 days.**

**Ex. 3.**

```{r}
sum(ga$visitsTumblr[format(ga$date, "%Y") == "2012"]) # 98808
```

**Ans.: 98808.**

**Ex. 4.**

```{r}
ga$dow <- weekdays(ga$date)
sort(tapply(ga$visitsTumblr, ga$dow, sum), decreasing = TRUE)
```

**Ans.: Monday — 25440 visits.**

**Ex. 5.**

```{r}
ga$ym <- format(ga$date, "%Y-%m")
sort(tapply(ga$visitsTumblr, ga$ym, sum), decreasing = TRUE)[1]
```

**Ans.: 2012-07 (17 484).**

**Ex. 6.**

```{r}
ga[which.max(ga$visitsTumblr), c("date", "visitsTumblr")]
```

**Ans.: 2011-11-28 — 4997 visits.**

**Ex. 7.**

```{r}
table(format(ga$date, "%Y")) # 2011: 365 ; 2012: 235
```

**Ans.: 2011 — 365 days, 2012 — 235 days.**

**Ex. 8.** → **`format(d, "%Y")`** (returns text; wrap in `as.numeric()` to get a number).

**Ex. 9.** → **returns the name of the day of the week.**

**Ex. 10.** → **`pdf("f.pdf")` → code → `dev.off()`** (without `dev.off()` the file won't be saved).

**Ex. 11.** → **shows the call stack after an error.**

**Ex. 12.** → **both work** (`substr` on the text representation, `str_extract` with regex).
