---
title: "09 — R Markdown"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

## Document structure

````
---
title: "My Report"
author: "Name"
date: "`r Sys.Date()`"
output: html_document   # or pdf_document, word_document
---

Text goes here. Markdown formatting applies.

```{r chunk-name, echo=TRUE, eval=TRUE}
# R code here
mean(1:10)
```
````

## Chunk options

| Option       | Default | Effect                                    |
|--------------|---------|-------------------------------------------|
| `echo`       | TRUE    | show source code in output                |
| `eval`       | TRUE    | run the code                              |
| `include`    | TRUE    | include chunk output in document          |
| `message`    | TRUE    | show messages (e.g. from library())       |
| `warning`    | TRUE    | show warnings                             |
| `fig.height` | 7       | figure height in inches                   |
| `fig.width`  | 7       | figure width in inches                    |
| `cache`      | FALSE   | cache results (skip re-running if unchanged) |

```{r eval=FALSE}
# Global options (in setup chunk)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Inline R

```
The mean mpg is `r mean(mtcars$mpg)` miles per gallon.
```
Renders as: The mean mpg is `r mean(mtcars$mpg)` miles per gallon.

## Markdown formatting

```markdown
# H1  ## H2  ### H3

**bold**   *italic*   `inline code`   ~~strikethrough~~

- bullet
  - nested bullet

1. numbered
2. list

[link text](https://url.com)

| Col1 | Col2 |
|------|------|
| a    | b    |
```

## Output formats

```yaml
output:
  html_document:
    toc: true          # table of contents
    toc_float: true    # floating TOC
    theme: flatly
  pdf_document:
    toc: true
```

## Knitting

```{r eval=FALSE}
rmarkdown::render("report.Rmd") # to HTML (default)
rmarkdown::render("report.Rmd", output_format = "pdf_document")
```

## Working directory in Rmd

The working directory for an Rmd file is **the directory containing the file**.
Use relative paths from there:

```{r eval=FALSE}
df <- read.csv("data/myfile.csv") # relative to Rmd location
```

## FlexDashboard skeleton

````
---
title: "Dashboard"
output: flexdashboard::flex_dashboard
---

Column
-------

### Chart A
```{r}
plot(1:10)
```

### Chart B
```{r}
hist(rnorm(100))
```
````
