---
title: "Sheet 02 — wide → long format (pivot_longer)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 02 — wide → long format (pivot_longer)

**Data:** `data/Czas reakcji.csv`. **Time:** ~35 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> This is an instructor-style sheet: load the data, **reshape the format**, then compute.
> The file is in wide format: two measurement columns (`Reakcja_przed`, `Reakcja_po`) for
> the same person.

**Ex. 1.** Load `data/Czas reakcji.csv`. How many people were studied? What is the mean
reaction time before and after (separately)?

```{r}
```

**Ex. 2.** Transform the data from **wide to long** format: instead of columns
`Reakcja_przed` and `Reakcja_po`, create a column `moment` (before/after) and a column `czas`.
Show the first rows.

```{r}
```

**Ex. 3.** On the long-format data, compute the mean reaction time at each moment
(group by `moment`). Does the result match Ex. 1?

```{r}
```

**Ex. 4.** How many people **improved** (time after < time before)? What was the mean
difference (`after − before`)? What was the percentage of people who improved and the
largest worsening?

```{r}
```

**Ex. 5.** From the long data, go back to wide format using `pivot_wider`. Verify that you
recovered the original columns.

```{r}
```

**Ex. 6.** Add a column `roznica = Reakcja_po − Reakcja_przed`, sort descending by
worsening and show the 3 people with the largest worsening.

```{r}
```

**Ex. 7.** Make a "before–after" plot: long data, `ggplot` with `moment` on the x-axis,
`czas` on the y-axis, connect measurements of the same person with a line (`group = id`).

```{r}
```

**Ex. 8.** (closed) Which `tidyr` function converts from wide to long format?

- `pivot_wider()`
- `pivot_longer()`
- `separate()`
- `unite()`

**Ex. 9.** (closed) To split `"PL-KRK"` into columns `kraj` and `miasto`:

- `unite(df, kraj, miasto, sep="-")`
- `separate(df, kol, into=c("kraj","miasto"), sep="-")`
- `str_split(df$kol)`

**Ex. 10.** (closed) A dumbbell chart (`geom_dumbbell` from `ggalt`) requires data in
**which** format?

- long
- wide
- nested

**Ex. 11.** (closed) `pivot_wider()` applied to the data from Ex. 2:

- returns to wide format
- gives the same as pivot_longer
- returns an error
