---
title: "Sheet 02 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 02 — solutions

**Ex. 1.**
```{r}
cr <- read.csv2("data/Czas reakcji.csv")
nrow(cr) # 25
colMeans(cr) # Reakcja_przed 0.2012, Reakcja_po 0.2688
```
**Ans.: 25 people; before = 0.2012; after = 0.2688.**

**Ex. 2.**
```{r}
library(dplyr)
library(tidyr)
cr_long <- cr %>%
  mutate(id = row_number()) %>%
  pivot_longer(
    cols = c(Reakcja_przed, Reakcja_po),
    names_to = "moment", values_to = "czas"
  )
head(cr_long)
```

**Ex. 3.**
```{r}
cr_long %>%
  group_by(moment) %>%
  summarise(srednia = mean(czas))
# po 0.2688 ; przed 0.2012  -> MATCHES
```

**Ex. 4.**
```{r}
sum(cr$Reakcja_po < cr$Reakcja_przed) # 4 people
mean(cr$Reakcja_po - cr$Reakcja_przed) # 0.0676
mean(cr$Reakcja_po < cr$Reakcja_przed) * 100 # 16 %
max(cr$Reakcja_po - cr$Reakcja_przed) # 0.16
```
**Ans.: improvement in 4 people (16%); mean difference +0.0676 (reaction time INCREASED on average);
largest worsening +0.16.**

**Ex. 5.**
```{r}
cr_wide <- cr_long %>% pivot_wider(names_from = moment, values_from = czas)
# columns: id, Reakcja_przed, Reakcja_po (same as original)
```

**Ex. 6.**
```{r}
cr %>%
  mutate(roznica = Reakcja_po - Reakcja_przed) %>%
  arrange(desc(roznica)) %>%
  head(3)
```

**Ex. 7.**
```{r}
library(ggplot2)
ggplot(cr_long, aes(x = moment, y = czas, group = id)) +
  geom_line(alpha = 0.4) +
  geom_point()
```

**Ex. 8.** → **`pivot_longer()`** (`gather` = longer, `spread` = wider — older names).

**Ex. 9.** → **`separate(df, kol, into=c("kraj","miasto"), sep="-")`**.

**Ex. 10.** → **wide** — `geom_dumbbell` connects two value columns
(before/after) in one row.

**Ex. 11.** → **returns to wide format** (`pivot_wider` is the inverse of `pivot_longer`).
