---
title: "Sheet 08 — Excel (xlsx), boxplot, groups, theory"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 08 — Excel (xlsx), boxplot, groups, theory

**Data:** `data/Prochnica.csv` (separator `;`), `data/Zawaly.xlsx`. **Time:** ~40 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

**Ex. 1.** Load `data/Prochnica.csv`. How many levels does the variable `Substancja_słodząca`
have and how many observations per level?

```{r}
```

**Ex. 2.** Draw a boxplot of `Próchnica` by `Substancja_słodząca`. Which substance has
the highest **mean** caries score?

```{r}
```

**Ex. 3.** Load `data/Zawaly.xlsx` (package `readxl`). How many rows and columns are there?

```{r}
```

**Ex. 4.** Compute the mean `BMI` separately for women and men (column `Plec`).

```{r}
```

**Ex. 5.** Do smokers have a higher mean `Cholesterol calkowity` than non-smokers
(column `Palenie`)?

```{r}
```

**Ex. 6.** Which substance in `Prochnica.csv` gives the most **variable** results
(highest standard deviation)?

```{r}
```

**Ex. 7.** Columns `HDL`, `LDL`, `Tg` in `Zawaly` form a lipid panel (wide format).
Transform them to **long** format (`frakcja`, `wartosc`) and compute the mean of each fraction.

```{r}
```

**Ex. 8.** How many patients belong to each `Rodzaj zawalu`?

```{r}
```

**Ex. 9.** (closed) Which function loads a `.xlsx` file?

- `read.csv("p.xlsx")`
- `readxl::read_excel("p.xlsx")`
- `load("p.xlsx")`

**Ex. 10.** (closed) A column whose name contains a space (`Cholesterol calkowity`)
in dplyr must be wrapped in:

- double quotes `"..."`
- backticks `` `...` ``
- parentheses `(...)`

**Ex. 11.** (closed) In `testthat`, checking that a function **throws an error** uses:

- `expect_equal()`
- `expect_true()`
- `expect_error()`

**Ex. 12.** (closed) To load the specific sheet "Dane" from an xlsx file:

- `read_excel("p.xlsx", tab="Dane")`
- `read_excel("p.xlsx", sheet="Dane")`
- `read_excel("p.xlsx", page="Dane")`
