---
title: "Sheet 08 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 08 — solutions

**Ex. 1.**

```{r}
pr <- read.csv2("data/Prochnica.csv")
table(pr$Substancja_słodząca)
```

**Ans.: 4 levels (Glicyna, Glukoza, Sacharoza, Sorbitol) with 12 observations each.**

**Ex. 2.**

```{r}
boxplot(Próchnica ~ Substancja_słodząca, data = pr)
tapply(pr$Próchnica, pr$Substancja_słodząca, mean)
# Glicyna 20.18; Glukoza 13.01; Sacharoza 38.70; Sorbitol 13.09
```

**Ans.: Sacharoza (≈ 38.7).**

**Ex. 3.**

```{r}
library(readxl)
za <- read_excel("data/Zawaly.xlsx")
dim(za) # 239 15
```

**Ans.: 239 rows, 15 columns.**

**Ex. 4.**

```{r}
library(dplyr)
za %>%
  group_by(Plec) %>%
  summarise(srednie_BMI = mean(BMI, na.rm = TRUE))
```

**Ans.: Female ≈ 26.5; Male ≈ 27.2.**

**Ex. 5.**

```{r}
za %>%
  group_by(Palenie) %>%
  summarise(chol = mean(`Cholesterol calkowity`, na.rm = TRUE))
```

**Ans.: smokers ≈ 243, non-smokers ≈ 245 → smokers have SLIGHTLY LOWER total cholesterol (minimal difference).**

**Ex. 6.**

```{r}
tapply(pr$Próchnica, pr$Substancja_słodząca, sd)
# Glicyna 2.73; Glukoza 2.85; Sacharoza 4.80; Sorbitol 2.62
```

**Ans.: Sacharoza (sd ≈ 4.80).**

**Ex. 7.**

```{r}
library(tidyr)
za %>%
  select(HDL, LDL, Tg) %>%
  pivot_longer(everything(), names_to = "frakcja", values_to = "wartosc") %>%
  group_by(frakcja) %>%
  summarise(srednia = mean(wartosc, na.rm = TRUE))
```

**Ans.: HDL ≈ 44.8; LDL ≈ 166; Tg ≈ 168.**

**Ex. 8.**

```{r}
table(za[["Rodzaj zawalu"]])
```

**Ans.: no infarction 103; non-transmural 51; transmural 85.**

**Ex. 9.** → **`readxl::read_excel("p.xlsx")`**.

**Ex. 10.** → **backticks** `` `Cholesterol calkowity` ``.

**Ex. 11.** → **`expect_error()`**.

**Ex. 12.** → **`read_excel("p.xlsx", sheet = "Dane")`**.
