---
title: "Sheet 03 — joining data frames (join), sorting, filtering, NA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 03 — joining data frames (join), sorting, filtering, NA

**Data:** `data/outcome-of-care-measures.csv`, `data/hospital-data.csv`. **Time:** ~40 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> American-locale files → use `read.csv()`. Mortality columns contain the text
> `"Not Available"` → after `as.numeric()` these become `NA`. Column 11 = mortality after
> heart attack (Heart Attack).

**Ex. 1.** Load both files. Join them on key `Provider.Number` (inner join). How many
records match?

```{r}
```

**Ex. 2.** Convert column 11 to numeric. How many missing values (`NA`) are created and why?

```{r}
```

**Ex. 3.** For state **TX**, find the hospital with the **lowest** mortality after heart attack.
What is its name and value? (Exclude `NA`.)

```{r}
```

**Ex. 4.** Which state has the most hospitals in the `outcome` dataset? Give the top 3.

```{r}
```

**Ex. 5.** What is the **national** mean heart attack mortality (excluding `NA`)?

```{r}
```

**Ex. 6.** Which hospital has the **highest** (worst) heart attack mortality and in which
state is it?

```{r}
```

**Ex. 7.** Which hospital is the **second** best (2nd lowest value) in TX?

```{r}
```

**Ex. 8.** How many hospitals in TX have heart attack mortality `<= 14`?

```{r}
```

**Ex. 9.** (closed) Which dplyr function keeps only records matching in BOTH data frames?

- `left_join()`
- `full_join()`
- `inner_join()`
- `bind_rows()`

**Ex. 10.** (closed) How to count how many values in vector `x` are missing?

- `length(x == NA)`
- `sum(is.na(x))`
- `count(NA, x)`

**Ex. 11.** (closed) An example of a **vector** graphics device is:

- PNG
- JPEG
- PDF
- TIFF

**Ex. 12.** (closed) After `initial_split()`, you extract the training set with:

- `train()`
- `training()`
- `get_train()`
- `analysis_set()`

**Ex. 13.** (closed) In this dataset (every `out` record has a match in `hos`),
`left_join(out, hos)` vs `inner_join(out, hos)`:

- will give different row counts
- will give the same row count
- left_join will error
