---
title: "Sheet 03 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 03 — solutions

```{r}
out <- read.csv("data/outcome-of-care-measures.csv", stringsAsFactors = FALSE)
hos <- read.csv("data/hospital-data.csv", stringsAsFactors = FALSE)
out$ha <- as.numeric(out[[11]]) # heart attack mortality (with NA warning)
library(dplyr)
```

**Ex. 1.**
```{r}
j <- inner_join(out, hos, by = "Provider.Number")
nrow(j) # 4706
```
**Ans.: 4706.**

**Ex. 2.**
```{r}
sum(is.na(out$ha)) # 1986
```
**Ans.: 1986 `NA`** — those hospitals had the text `"Not Available"` in the column.

**Ex. 3.**
```{r}
out %>%
  filter(State == "TX", !is.na(ha)) %>%
  arrange(ha, Hospital.Name) %>%
  slice(1) %>%
  select(Hospital.Name, ha)
```
**Ans.: CYPRESS FAIRBANKS MEDICAL CENTER, value 12.**

**Ex. 4.**
```{r}
sort(table(out$State), decreasing = TRUE)[1:3]
```
**Ans.: TX (370), CA (341), NY (185).**

**Ex. 5.**
```{r}
mean(out$ha, na.rm = TRUE) # 15.449
```
**Ans.: ≈ 15.45.**

**Ex. 6.**
```{r}
out %>%
  filter(!is.na(ha)) %>%
  arrange(desc(ha)) %>%
  slice(1) %>%
  select(Hospital.Name, State, ha)
```
**Ans.: MEDICAL CENTER SOUTH ARKANSAS (AR), ha = 21.9.**

**Ex. 7.**
```{r}
out %>%
  filter(State == "TX", !is.na(ha)) %>%
  arrange(ha, Hospital.Name) %>%
  slice(2)
```
**Ans.: HOUSTON NORTHWEST MEDICAL CENTER (ha = 12.4).**

**Ex. 8.**
```{r}
sum(out$ha[out$State == "TX"] <= 14, na.rm = TRUE) # 21
```
**Ans.: 21.**

**Ex. 9.** → **`inner_join()`**.

**Ex. 10.** → **`sum(is.na(x))`**.

**Ex. 11.** → **PDF** (PNG/JPEG/TIFF are bitmaps).

**Ex. 12.** → **`training()`** (test set: `testing()`).

**Ex. 13.** → **will give the same row count** (all `out` records have a match in `hos`).
