---
title: "04 — Strings, Regex, Dates"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
library(stringr)
```

## stringr core functions

```{r}
x <- c("Smith, John", "Obama, Barack", "United States", "united kingdom")

str_detect(x, "^United") # TRUE/FALSE vector
str_detect(x, regex("united", ignore_case = TRUE))
str_extract(x, "[A-Z][a-z]+") # first match per element
str_extract_all(x, "[A-Z][a-z]+") # all matches (returns list)
str_replace(x, ",.*", "") # remove from comma onward
str_replace_all(x, "[, ]", "_") # replace every match
str_split("a;b;c", ";")
str_trim("  abc  ") # strip whitespace
str_sub("2024-01-15", 1, 4) # characters 1–4
str_length(x)
```

## Base R equivalents

```{r}
grep("^United", x) # indices of matches
grep("^United", x, value = TRUE) # matching values
grepl("^United", x) # TRUE/FALSE vector
gsub("[, ]", "_", x) # replace all (= str_replace_all)
sub("[, ]", "_", x) # replace first only
```

## Regex patterns

| Pattern      | Meaning                              |
|--------------|--------------------------------------|
| `^abc`       | starts with "abc"                    |
| `abc$`       | ends with "abc"                      |
| `[0-9]+`     | one or more digits                   |
| `[0-9]{4}`   | exactly 4 digits                     |
| `[^,]+`      | one or more non-comma characters     |
| `.`          | any single character (except `\n`)   |
| `(a\|b)`     | a or b                               |
| `\\d`        | digit (= `[0-9]`)                    |
| `\\s`        | whitespace                           |

## Common string tasks

```{r}
# Parse "Last, First" → first name
names <- c("Smith, John", "Obama, Barack")
str_extract(names, "(?<=, ).+") # lookbehind
str_trim(str_replace(names, "^[^,]+, ", "")) # alternative
```

```{r}
# Count countries starting with "United"
countries <- c("United States", "Germany", "United Kingdom", "United Arab Emirates")
sum(str_detect(countries, "^United")) # 3
length(grep("^United", countries)) # 3
```

```{r}
# Remove commas and spaces from numeric strings
amounts <- c("28,750,956", "1 234 567")
as.numeric(gsub("[ ,]", "", amounts))
```

---

## Dates — base R

```{r}
d <- as.Date("2012-03-15")
class(d)
```

```{r}
# Parsing different formats
as.Date("15/03/2012", format = "%d/%m/%Y")
as.Date("4/18/1950 0:00:00", format = "%m/%d/%Y %H:%M:%S")
```

```{r}
# Extracting components
format(d, "%Y") # year as string "2012"
format(d, "%m") # month "03"
format(d, "%Y-%m") # year-month "2012-03"
as.integer(format(d, "%Y")) # year as number
weekdays(d) # "Thursday" (locale-dependent; Polish: "czwartek")
months(d)
```

```{r}
# Arithmetic
d2 <- as.Date("2012-12-31")
as.integer(d2 - d) # days between
d + 30 # date plus 30 days
```

```{r}
# Typical exam pattern: filter by year, count, aggregate
ga <- read.csv("../../sheets/custom/06-daty/data/gaData.csv")
ga$date <- as.Date(ga$date)
sum(ga$visitsTumblr[format(ga$date, "%Y") == "2012"])
table(format(ga$date, "%Y"))
```

## lubridate

```{r message=FALSE}
library(lubridate)
d <- ymd("2012-03-15")
dmy("15/03/2012")
mdy("03/15/2012")

year(d)
month(d)
day(d)
wday(d, label = TRUE) # labelled weekday
today()
now()
```

## format() codes

| Code | Meaning         | Example |
|------|-----------------|---------|
| `%Y` | 4-digit year    | 2012    |
| `%m` | month 01–12     | 03      |
| `%d` | day 01–31       | 15      |
| `%H` | hour 00–23      | 08      |
| `%M` | minute 00–59    | 30      |
| `%A` | full day name   | Thursday |
