---
title: "Sheet 12 — tidymodels (ML): olive oil classification"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 12 — tidymodels (ML): olive oil classification

**Data:** `data/olive.rda` (object `olive`). **Time:** ~35 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> `olive`: 572 olive oil samples, composition of 8 fatty acids + `Region` (1/2/3) and `Area` (1–9).
> Goal: train a model to identify **Region** from fatty acid composition.
> Note: `Area` is derived from Region (it "cheats") — remove it from the predictors.
> Loading: `load("data/olive.rda")` creates the object `olive` (NOT `x <- load(...)`).

## Theory section (closed — these are the most frequent instructor questions)

**T1.** Function that splits data into training and test sets:

- `split()`
- `train_test_split()`
- `initial_split()`
- `data_partition()`

**T2.** After `initial_split()`, you extract the sets with:

- `train()` / `test()`
- `training()` / `testing()`
- `get_train()` / `get_test()`

**T3.** Feature preparation/transformation (e.g. normalization) in tidymodels:

- `recipe()` + `step_*()`
- `prep_data()`
- `mutate()`
- `transform()`

**T4.** The model definition (type + engine + mode) is created in the package:

- `parsnip`
- `ggplot2`
- `dplyr`
- `caret`

**T5.** Combining a recipe and a model into one object:

- `pipeline()`
- `workflow()`
- `bind_rows()`
- `merge()`

**T6.** Training the model / predicting:

- `fit()` / `predict()`
- `train()` / `guess()`
- `learn()` / `apply()`

**T7.** Evaluating quality (accuracy, rmse, roc_auc) — package:

- `yardstick`
- `metrics2`
- `scales`
- `broom`

**T8.** k-fold cross-validation:

- `vfold_cv()`
- `crossval()`
- `kfold()`
- `cv_split()`

## Practical section (run the code)

**P1.** Load `data/olive.rda`. Convert `Region` to `factor`, remove column `Area`.

```{r}
```

**P2.** Split the data 75/25 with stratification on `Region` (set `set.seed(123)`).

```{r}
```

**P3.** Build a recipe: `Region ~ .` + normalize the numeric predictors.

```{r}
```

**P4.** Define a decision tree model (engine `rpart`, classification mode), combine in a
`workflow`, train it, and compute **accuracy** on the test set.

```{r}
```
