---
title: "Sheet 12 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 12 — solutions

## Theory

- **T1** → `initial_split()`
- **T2** → `training()` / `testing()`
- **T3** → `recipe()` + `step_*()`
- **T4** → `parsnip`
- **T5** → `workflow()`
- **T6** → `fit()` / `predict()`
- **T7** → `yardstick`
- **T8** → `vfold_cv()`

## Practice — full working code

```{r}
library(tidymodels) # or: rsample, recipes, parsnip, workflows, yardstick
load("data/olive.rda")

# P1 — preparation
ol <- olive %>%
  mutate(Region = factor(Region)) %>%
  select(-Area)

# P2 — train/test split (stratified by class)
set.seed(123)
podzial <- initial_split(ol, prop = 0.75, strata = Region)
treningowe <- training(podzial) # 428 obs.
testowe <- testing(podzial) # 144 obs.

# P3 — recipe: formula + processing steps
przepis <- recipe(Region ~ ., data = treningowe) %>%
  step_normalize(all_numeric_predictors())

# P4 — model + workflow + training + evaluation
model <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("classification")
wf <- workflow() %>%
  add_recipe(przepis) %>%
  add_model(model)
fit_wf <- fit(wf, data = treningowe)

pred <- predict(fit_wf, new_data = testowe) %>%
  bind_cols(testowe %>% select(Region))

accuracy(pred, truth = Region, estimate = .pred_class) # accuracy = 1.00
conf_mat(pred, truth = Region, estimate = .pred_class) # perfect confusion matrix

# (optional) cross-validation:
set.seed(1)
folds <- vfold_cv(treningowe, v = 5, strata = Region)
```

**Result: accuracy ≈ 1.00** (fatty acids separate the regions very well).
Other models instead of `decision_tree()`: `rand_forest()` (engine `"ranger"`),
`multinom_reg()` / `logistic_reg()`, `nearest_neighbor()`.

## Variant WITHOUT tidymodels (if packages are unavailable)

```{r}
load("data/olive.rda")
ol <- olive
ol$Region <- factor(ol$Region)
ol$Area <- NULL
set.seed(123)
idx <- sample(nrow(ol), round(0.75 * nrow(ol)))
train <- ol[idx, ]
test <- ol[-idx, ]

library(rpart) # decision tree
m <- rpart(Region ~ ., data = train, method = "class")
pred <- predict(m, test, type = "class")
mean(pred == test$Region) # ≈ 0.986
table(pred, test$Region)

library(class) # k-NN
knn(scale(train[, -1]), scale(test[, -1]), train$Region, k = 5) # ≈ 1.00
```

## tidymodels pipeline (memorise the order!)

`initial_split` → `training`/`testing` → `recipe` + `step_*` → model (`parsnip`) →
`workflow` (`add_recipe` + `add_model`) → `fit` → `predict` → `metrics`/`accuracy`.
