---
title: "Sheet 05 — missing values (NA), grouping, type theory"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 05 — missing values (NA), grouping, type theory

**Data:** `data/q1_data.csv` (`airquality` dataset: Ozone, Solar.R, Wind, Temp, Month, Day).
**Time:** ~35 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

**Ex. 1.** Load `data/q1_data.csv`. Compute the mean ozone concentration (`Ozone`). Why
does plain `mean(Ozone)` return `NA` and how do you fix it? How many missing values are
in `Ozone`?

```{r}
```

**Ex. 2.** Compute the mean temperature (`Temp`) in each month (`Month`).

```{r}
```

**Ex. 3.** In which month was the mean ozone concentration highest (excluding `NA`)?

```{r}
```

**Ex. 4.** What is the correlation between `Ozone` and `Temp` (on complete observations)?

```{r}
```

**Ex. 5.** How many rows are **complete** (no `NA` at all)? Out of how many total?

```{r}
```

**Ex. 6.** What is the mean `Ozone` on warm days (`Temp > 80`, excluding `NA`)?

```{r}
```

**Ex. 7.** Compute the correlation matrix for `Ozone, Solar.R, Wind, Temp` (on complete
observations). Which pair is most strongly correlated **positively**, and which
**negatively**?

```{r}
```

**Ex. 8.** (closed) What is the class of `x <- c(4, TRUE)`?

- logical
- character
- numeric
- list

**Ex. 9.** (closed) What does `x[[2]]` give for `x <- list(2, "a", "b", TRUE)`?

- a list with "a" and "b"
- a character vector with the letter "a"
- the number 2

**Ex. 10.** (closed) What will `rbind(c(1,3,5), c(3,2,10))` return?

- a vector of length 3
- a 3×2 matrix
- a 2×3 matrix

**Ex. 11.** (closed) Default fitting method in `geom_smooth()` (without `method`)?

- lm
- loess
- glm
- gam

**Ex. 12.** (closed) To make `cor()` compute despite `NA`, you use:

- `cor(x, y, na.rm = TRUE)`
- `cor(x, y, use = "complete.obs")`
- `cor(na.omit(x), y)`
