---
title: "Sheet 05 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 05 — solutions

**Ex. 1.**

```{r}
aq <- read.csv("data/q1_data.csv")
mean(aq$Ozone) # NA — any arithmetic with NA gives NA
mean(aq$Ozone, na.rm = TRUE) # 42.1293
sum(is.na(aq$Ozone)) # 37
```

**Ans.: `mean` returns `NA` because the column contains missing values; use `na.rm = TRUE`. Mean
≈ 42.13; 37 missing values.**

**Ex. 2.**

```{r}
library(dplyr)
aq %>%
  group_by(Month) %>%
  summarise(srednia_temp = mean(Temp))
# base: tapply(aq$Temp, aq$Month, mean)
```

**Ans.: May 65.6; Jun 79.1; Jul 83.9; Aug 84.0; Sep 76.9.**

**Ex. 3.**

```{r}
aq %>%
  group_by(Month) %>%
  summarise(o = mean(Ozone, na.rm = TRUE)) %>%
  arrange(desc(o)) %>%
  head(1)
```

**Ans.: month 8 (August), ≈ 60.0.**

**Ex. 4.**

```{r}
cor(aq$Ozone, aq$Temp, use = "complete.obs") # 0.6984
```

**Ans.: ≈ 0.70.**

**Ex. 5.**

```{r}
sum(complete.cases(aq)) # 111
nrow(aq) # 153
```

**Ans.: 111 of 153.**

**Ex. 6.**

```{r}
mean(aq$Ozone[aq$Temp > 80], na.rm = TRUE) # 64.148
```

**Ans.: ≈ 64.15.**

**Ex. 7.**

```{r}
cor(aq[, c("Ozone", "Solar.R", "Wind", "Temp")], use = "complete.obs")
```

**Ans.: strongest POSITIVE pair: Ozone–Temp (0.699); strongest NEGATIVE pair: Ozone–Wind (−0.612).**

**Ex. 8.** → **numeric** (`TRUE` → `1`).

**Ex. 9.** → **a character vector with the letter "a"** (`[[ ]]` extracts a single element).

**Ex. 10.** → **a 2×3 matrix** (each vector becomes a row).

**Ex. 11.** → **loess**.

**Ex. 12.** → **`cor(x, y, use = "complete.obs")`** (`cor` has no `na.rm` argument).
