---
title: "Sheet 14 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Ex. 1.**

```{r}
storm <- read.csv("data/data_StormData.csv.bz2")
dim(storm) # 902297 × 37
length(unique(storm$EVTYPE)) # ~985
```

**Ans.: 902,297 rows, ~985 unique EVTYPE (lots of inconsistent naming).**

**Ex. 2.**

```{r}
library(dplyr)
storm$year <- as.integer(format(as.Date(storm$BGN_DATE, "%m/%d/%Y %H:%M:%S"), "%Y"))
storm %>%
  count(year) %>%
  arrange(desc(n)) %>%
  slice(1:3)
```

**Ans.: Years ~2011, 2010, 2009 have the most events (registration increased over time).**

**Ex. 3.**

```{r}
storm %>%
  group_by(EVTYPE) %>%
  summarise(fatal = sum(FATALITIES, na.rm = TRUE)) %>%
  arrange(desc(fatal)) %>%
  slice(1:5)
```

**Ans.: TORNADO — 5,633 fatalities.**

**Ex. 4.**

```{r}
storm %>%
  group_by(EVTYPE) %>%
  summarise(inj = sum(INJURIES, na.rm = TRUE)) %>%
  arrange(desc(inj)) %>%
  slice(1:5)
```

**Ans.: TORNADO — 91,346 injuries.**

**Ex. 5.**

```{r}
storm %>%
  group_by(EVTYPE) %>%
  summarise(total_health = sum(FATALITIES + INJURIES, na.rm = TRUE)) %>%
  arrange(desc(total_health)) %>%
  slice(1:5)
```

**Ans.: TORNADO = 96,979 (combined), far ahead of EXCESSIVE HEAT (8,428).**

**Ex. 6.**

```{r}
exp_mult <- function(e) {
  e <- toupper(trimws(as.character(e)))
  ifelse(e == "K", 1e3,
    ifelse(e == "M", 1e6,
      ifelse(e == "B", 1e9,
        ifelse(grepl("^[0-9]$", e), 10^as.numeric(e), 1)
      )
    )
  )
}

storm$prop_dmg <- storm$PROPDMG * exp_mult(storm$PROPDMGEXP)
# Check:
sum(storm$prop_dmg > 0)
```

**Key point:** `toupper()` + `trimws()` because the data contains "m", "k", spaces.**

**Ex. 7.**

```{r}
storm$crop_dmg <- storm$CROPDMG * exp_mult(storm$CROPDMGEXP)
storm$econ_dmg <- storm$prop_dmg + storm$crop_dmg

storm %>%
  group_by(EVTYPE) %>%
  summarise(total_econ = sum(econ_dmg, na.rm = TRUE)) %>%
  arrange(desc(total_econ)) %>%
  slice(1:5)
```

**Ans.: FLOOD ~$150 B, HURRICANE/TYPHOON ~$72 B, TORNADO ~$57 B.**

**Ex. 8.**

```{r}
storm %>%
  filter(EVTYPE %in% c("TORNADO", "FLOOD")) %>%
  group_by(EVTYPE) %>%
  summarise(
    zdrowie = sum(FATALITIES + INJURIES, na.rm = TRUE),
    ekonomia = sum(econ_dmg, na.rm = TRUE)
  )
```

**Ans.: TORNADO wins on health (96,979 vs 7,259); FLOOD wins on economics (~$150 B vs ~$57 B).**

**Ex. 9.**

```{r}
storm %>%
  filter(EVTYPE == "TORNADO") %>%
  group_by(STATE) %>%
  summarise(fatal = sum(FATALITIES, na.rm = TRUE)) %>%
  arrange(desc(fatal)) %>%
  slice(1:3)
```

**Ans.: Alabama (AL) or Texas (TX) — depends on the data scope.**

**Ex. 10.**

```{r}
storm %>%
  group_by(EVTYPE) %>%
  summarise(
    n_events = n(),
    fatal    = sum(FATALITIES, na.rm = TRUE),
    rate     = fatal / n_events
  ) %>%
  filter(n_events >= 10) %>%
  arrange(desc(rate)) %>%
  slice(1:5)
```

**Ans.: Highest rate belongs to rare but deadly events (e.g. TORNADOES, TSTM WIND F2).**

**Ex. 11.** → **`read.csv("file.csv.bz2")` — R unpacks `.bz2` automatically.**

**Ex. 12.** → **billions — multiplier 1,000,000,000.**

**Ex. 13.** → **EVTYPE starting with "TORNADO"** (`^` anchors to the start).

**Ex. 14.**

```{r}
storm %>%
  filter(year >= 1990, year <= 2010) %>%
  group_by(year) %>%
  summarise(total_econ = sum(econ_dmg, na.rm = TRUE)) %>%
  arrange(desc(total_econ))
```

**Ans.: Visible increase in later years; 2005 (Katrina) or 2006 often appears as the costliest year.**
