---
title: "Sheet 09 — TWO datasets: daily PM2.5 measurements + station metadata"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 09 — TWO datasets: daily PM2.5 measurements + station metadata

**Data:** `data/daily_SPEC_2014.csv.bz2` (measurements), `data/aqs_sites.xlsx` (station metadata).
**Time:** ~50 min. **Level: hard.**

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> Reconstructed "daily measurements quiz". Load measurements with `readr::read_csv`
> (R unpacks `.bz2` automatically), metadata with `readxl::read_excel`. In all means
> ignore missing values (`na.rm = TRUE`).
>
> **Key pitfall:** in the measurements, `State Code`/`County Code`/`Site Num` are read as
> text with leading zeros ("06", "081"), while in the metadata they are numbers. Before
> joining, harmonise types (e.g. `as.integer()` on both sides).

**Ex. 1.** What is the mean `Arithmetic Mean` for `"Bromine PM2.5 LC"` in the state of
**Wisconsin**?

- 0.003960
- 1.485826
- 0.515084
- 0.458098

```{r}
```

**Ex. 2.** Which component (`Parameter Name`) has the highest mean?

- Sulfur PM2.5 LC
- OC CSN Unadjusted PM2.5 LC TOT
- EC2 PM2.5 LC
- Sodium PM2.5 LC

```{r}
```

**Ex. 3.** Which location (State/County/Site) has the highest mean `"Sulfate PM2.5 LC"`?

- 48/203/0002
- 31/055/0019
- 10/001/0003
- 39/081/0017

```{r}
```

**Ex. 4.** Absolute difference between the means of `"EC PM2.5 LC TOR"` in California
vs Arizona?

- 0.018567
- 0.06320775
- 0.17648069
- 0.09476190

```{r}
```

**Ex. 5.** Median of `"OC PM2.5 LC TOR"` in the **west** (Longitude < −100)?

- 0.5175
- 0.4300
- 0.2600
- 0.0269

```{r}
```

**Ex. 6.** (2 datasets) How many stations have both `Land Use == "RESIDENTIAL"` and
`Location Setting == "SUBURBAN"`?

- 3527
- 1610
- 2233
- 1008

```{r}
```

**Ex. 7.** (2 datasets) Median of `"EC PM2.5 LC TOR"` among RESIDENTIAL+SUBURBAN stations
in the **east** (Longitude ≥ −100)?

- 0.6100
- 0.0460
- 0.5880
- 0.1740

```{r}
```

**Ex. 8.** (2 datasets) Among `Land Use == "COMMERCIAL"` stations, which **month** has
the highest mean `"Sulfate PM2.5 LC"`?

- November
- February
- August
- June

```{r}
```

**Ex. 9.** For monitor State 6 / County 65 / Site 8001: on how many **days** does the
sum of `"Sulfate PM2.5 LC"` and `"Total Nitrate PM2.5 LC"` exceed 10? (when a day has
multiple values for a component — average them first)

- 5
- 11
- 24
- 50

```{r}
```
