---
title: "Sheet 09 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 09 — solutions

```{r}
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
spec <- read_csv("data/daily_SPEC_2014.csv.bz2") # ~2.1 million rows
sites <- read_excel("data/aqs_sites.xlsx")
# integer join key (leading zeros "06" -> 6):
spec <- spec %>%
  mutate(
    skey = as.integer(`State Code`),
    ckey = as.integer(`County Code`),
    nkey = as.integer(`Site Num`)
  )
sites <- sites %>%
  mutate(
    skey = as.integer(`State Code`),
    ckey = as.integer(`County Code`),
    nkey = as.integer(`Site Number`)
  )
```

**Ex. 1.** → **0.003960**
```{r}
spec %>%
  filter(`State Name` == "Wisconsin", `Parameter Name` == "Bromine PM2.5 LC") %>%
  summarise(mean(`Arithmetic Mean`, na.rm = TRUE))
```

**Ex. 2.** → **OC CSN Unadjusted PM2.5 LC TOT** (≈ 67.8)
```{r}
params <- c(
  "Sulfur PM2.5 LC", "OC CSN Unadjusted PM2.5 LC TOT",
  "EC2 PM2.5 LC", "Sodium PM2.5 LC"
)
spec %>%
  filter(`Parameter Name` %in% params) %>%
  group_by(`Parameter Name`) %>%
  summarise(m = mean(`Arithmetic Mean`, na.rm = TRUE)) %>%
  arrange(desc(m))
```
(In the full dataset the highest values belong to meteorological parameters like pressure — that is why
we restrict to the components from the answer list.)

**Ex. 3.** → **39 / 081 / 0017** (≈ 3.18)
```{r}
spec %>%
  filter(`Parameter Name` == "Sulfate PM2.5 LC") %>%
  group_by(`State Code`, `County Code`, `Site Num`) %>%
  summarise(m = mean(`Arithmetic Mean`, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(m)) %>%
  head(1)
```

**Ex. 4.** → **0.018567** (CA 0.198 − AZ 0.179)
```{r}
q4 <- spec %>%
  filter(
    `Parameter Name` == "EC PM2.5 LC TOR",
    `State Name` %in% c("California", "Arizona")
  ) %>%
  group_by(`State Name`) %>%
  summarise(m = mean(`Arithmetic Mean`, na.rm = TRUE))
abs(diff(q4$m))
```

**Ex. 5.** → **0.4300**
```{r}
spec %>%
  filter(`Parameter Name` == "OC PM2.5 LC TOR", Longitude < -100) %>%
  summarise(median(`Arithmetic Mean`, na.rm = TRUE))
```

**Ex. 6.** → **3527**
```{r}
sites %>%
  filter(`Land Use` == "RESIDENTIAL", `Location Setting` == "SUBURBAN") %>%
  nrow()
```

**Ex. 7.** → **0.6100** (key step: **join** measurements with metadata)
```{r}
rs <- sites %>%
  filter(`Land Use` == "RESIDENTIAL", `Location Setting` == "SUBURBAN") %>%
  distinct(skey, ckey, nkey)
spec %>%
  semi_join(rs, by = c("skey", "ckey", "nkey")) %>%
  filter(`Parameter Name` == "EC PM2.5 LC TOR", Longitude >= -100) %>%
  summarise(median(`Arithmetic Mean`, na.rm = TRUE))
```
`semi_join` keeps rows of `spec` that have a match in `rs` (no extra columns added).

**Ex. 8.** → **February** (month 2; ≈ 2.02)
```{r}
com <- sites %>%
  filter(`Land Use` == "COMMERCIAL") %>%
  distinct(skey, ckey, nkey)
spec %>%
  semi_join(com, by = c("skey", "ckey", "nkey")) %>%
  filter(`Parameter Name` == "Sulfate PM2.5 LC") %>%
  mutate(miesiac = as.integer(format(`Date Local`, "%m"))) %>%
  group_by(miesiac) %>%
  summarise(m = mean(`Arithmetic Mean`, na.rm = TRUE)) %>%
  arrange(desc(m)) %>%
  head(1)
```

**Ex. 9.** → **11**
```{r}
q9 <- spec %>%
  filter(
    skey == 6, ckey == 65, nkey == 8001,
    `Parameter Name` %in% c("Sulfate PM2.5 LC", "Total Nitrate PM2.5 LC")
  ) %>%
  group_by(`Date Local`, `Parameter Name`) %>% # average duplicates within a day
  summarise(v = mean(`Arithmetic Mean`, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = `Parameter Name`, values_from = v) %>%
  mutate(suma = `Sulfate PM2.5 LC` + `Total Nitrate PM2.5 LC`)
sum(q9$suma > 10, na.rm = TRUE)
```
**Pitfall:** `filter(\`State Code\` == 6, ...)` will NOT work — `State Code` is text `"06"`.
Use the integer key `skey`/`ckey`/`nkey`.
