---
title: "Sheet 04 — stringr / regex + filtering and groups"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sheet 04 — stringr / regex + filtering and groups

**Data:** `data/cel_volden_wiseman.csv` (codebook: `data/cel_data_codebook.txt`).
**Time:** ~40 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> Dataset: US Congress members. Key columns: `thomas_name` (format `"Last, First"`),
> `congress`, `dem` (1 = Democrat), `female` (1 = woman), `st_name` (state),
> `les` (Legislative Effectiveness Score).

**Ex. 1.** Load the data. For `congress == 110`, compute the mean `les` separately for
Democrats and Republicans (`dem`).

```{r}
```

**Ex. 2.** Extract just the **last name** (part before the comma) from `thomas_name` into
a new column `nazwisko`. Use `stringr`.

```{r}
```

**Ex. 3.** How many Congress members in `congress == 110` have a last name starting with
the letter **"M"**?

```{r}
```

**Ex. 4.** How many women (`female == 1`) served in `congress == 110`?

```{r}
```

**Ex. 5.** For `congress == 110`, find the state (`st_name`) with the highest mean `les`
(consider only states with at least 5 observations).

```{r}
```

**Ex. 6.** How many distinct states (`st_name`) are represented in `congress == 110`?

```{r}
```

**Ex. 7.** How many Congress members in `congress == 110` have a `les` **above their
state's mean**? (Compute the state mean and compare to individual `les`.)

```{r}
```

**Ex. 8.** Who was the most effective (max `les`) Congress member in `congress == 110`?

```{r}
```

**Ex. 9.** Also extract the **first name** (part after the comma) from `thomas_name` into
column `imie` and trim the leading space.

```{r}
```

**Ex. 10.** In `congress == 110`, did women have a higher mean `les` than men?

```{r}
```

**Ex. 11.** (closed) Which `stringr` function checks whether a string **contains** a
pattern?

- `str_extract()`
- `str_detect()`
- `str_replace()`
- `str_sub()`

**Ex. 12.** (closed) Regular expression for strings starting with "United":

- `grep("United$", x)`
- `grep("*United", x)`
- `grep("^United", x)`

**Ex. 13.** (closed) `gsub("a","b",x)` as opposed to `sub("a","b",x)`:

- replaces only the first occurrence
- replaces all occurrences
- removes letters
