---
title: "Sheet 15 ‚ÄĒ hurricanes: read_fwf, wide‚Üílong, wind analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Data:** `data/ebtrk_atlc_1988_2015.txt` (11,824 rows, fixed-width format).
**Time:** ~45 min.

Set working directory to this folder. Solutions: `rozwiazania.Rmd`.

> A text file **without a header** in FWF (fixed-width format).
> Each row = one hurricane measurement (every 6 hours).
> Read with `readr::read_fwf()` and `fwf_widths(c(...), c("name", ...))`.
> Column widths and names are in the solutions.
> Key columns: `storm_id`, `storm_name`, `year`, `wind_kts` (wind speed in knots),
> 12 wind radius columns:
> `ts_radius_{NE,SE,SW,NW}` (‚Č•34 kt), `r50_{NE,SE,SW,NW}` (‚Č•50 kt), `r64_{NE,SE,SW,NW}` (‚Č•64 kt).
> Value -99 = missing data.

**Ex. 1.** Load the file using `read_fwf()` with `fwf_widths()`. Check `dim()` and `names()`.
How many unique hurricanes (`storm_id`) are in the dataset?

```{r}
```

**Ex. 2.** Convert columns `wind_kts`, `year`, `lat`, `lon` to `numeric` type.
Replace values of -99 in the wind radius columns with `NA`.

```{r}
```

**Ex. 3.** What is the year range in the data? How many observations (measurements) does
each year have ‚ÄĒ which year had the most hurricane activity?

```{r}
```

**Ex. 4.** What is the maximum wind speed in the dataset? Which hurricane (`storm_name`)
achieved it and in which year?

```{r}
```

**Ex. 5.** How many observations have wind speed ‚Č• 64 knots (hurricane force on the
Beaufort scale)? What percentage of all measurements does this represent?

```{r}
```

**Ex. 6.** Transform the data from wide to long format (`pivot_longer`) for the
12 wind radius columns (`ts_radius_NE` through `r64_NW`). The resulting data frame should
have columns `quadrant` and `radius`. Filter out rows with `radius <= 0` or `NA`.

```{r}
```

**Ex. 7.** After the transformation: how many rows remain after filtering out NA and ‚Č§ 0?

```{r}
```

**Ex. 8.** Split the `quadrant` column (e.g. "r50_NE") into `threshold` (e.g. "r50")
and `direction` (e.g. "NE"). Which direction (NE/SE/SW/NW) has the largest mean radius
for the 50-knot threshold (`r50`)?

```{r}
```

**Ex. 9.** For each year, compute the maximum wind speed across all measurements.
Which year had the strongest hurricane? What was its name?

```{r}
```

**Ex. 10.** Compute the mean wind speed per hurricane (`storm_id`). Which hurricane had
the highest **mean** (not maximum) wind speed over its entire lifetime?

```{r}
```

**Ex. 11.** (closed) `fwf_widths(c(7, 10, 6), c("id", "name", "date"))` specifies:

- the number of rows to skip
- the widths (in characters) of successive columns and their names
- column separators at positions 7, 10, 6
- data types of the columns

**Ex. 12.** (closed) `pivot_longer(cols = a:z, names_to = "variable", values_to = "value")` causes:

- column names a‚Ä¶z to become values in the column `variable`
- the column `variable` to contain numbers from 1 to 26
- columns a‚Ä¶z to be duplicated
- values to go into many new columns

**Ex. 13.** (closed) The value -99 in this database means:

- hurricane category -99
- missing measurement data
- 99 nautical miles in the westward direction
- GPS read error

**Ex. 14.** Classify each observation by wind speed:
"TS" (‚Č•34 kt), "Cat1" (‚Č•64 kt), "Cat3" (‚Č•96 kt), "Cat5" (‚Č•137 kt).
Count the number of observations in each category.

```{r}
```
