---
title: "Sheet 15 — solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Ex. 1.**

```{r}
library(readr)
hur <- read_fwf(
  "data/ebtrk_atlc_1988_2015.txt",
  fwf_widths(
    c(7, 10, 6, 5, 5, 6, 4, 4, 4, 4, 5, 5, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 2, 6),
    c(
      "storm_id", "storm_name", "date", "year", "lat", "lon",
      "wind_kts", "pressure",
      "ts_radius_NE", "ts_radius_SE", "ts_radius_SW", "ts_radius_NW",
      "r50_NE", "r50_SE", "r50_SW", "r50_NW",
      "r64_NE", "r64_SE", "r64_SW", "r64_NW",
      "r100_NE", "r100_SE", "r100_SW", "r100_NW",
      "ts_type", "distance"
    )
  ),
  col_types = cols(.default = col_character())
)
dim(hur) # 11824 × 26
length(unique(hur$storm_id)) # 377
```

**Ans.: 11,824 measurements, 377 unique hurricanes.**

**Ex. 2.**

```{r}
library(dplyr)

radius_cols <- c(
  "ts_radius_NE", "ts_radius_SE", "ts_radius_SW", "ts_radius_NW",
  "r50_NE", "r50_SE", "r50_SW", "r50_NW",
  "r64_NE", "r64_SE", "r64_SW", "r64_NW"
)

hur <- hur %>%
  mutate(
    wind_kts = as.numeric(wind_kts),
    year = as.integer(year),
    lat = as.numeric(lat),
    lon = as.numeric(lon),
    across(all_of(radius_cols), ~ ifelse(as.numeric(.) == -99, NA, as.numeric(.)))
  )
```

**Ex. 3.**

```{r}
range(hur$year, na.rm = TRUE) # 1988 2015

hur %>%
  count(year) %>%
  arrange(desc(n)) %>%
  slice(1:5)
```

**Ans.: Range 1988–2015. Most active years: usually years with many seasons (e.g. 2005).**

**Ex. 4.**

```{r}
hur %>%
  filter(wind_kts == max(wind_kts, na.rm = TRUE)) %>%
  select(storm_name, year, wind_kts)
```

**Ans.: GILBERT, 1988, 160 knots.**

**Ex. 5.**

```{r}
hur_force <- sum(hur$wind_kts >= 64, na.rm = TRUE) # 2984
hur_force
hur_force / nrow(hur) * 100 # ~25.2%
```

**Ans.: 2984 observations (≈25% of all measurements) had hurricane force.**

**Ex. 6.**

```{r}
library(tidyr)

long <- hur %>%
  select(storm_id, storm_name, year, date, wind_kts, all_of(radius_cols)) %>%
  pivot_longer(
    cols = all_of(radius_cols),
    names_to = "quadrant",
    values_to = "radius"
  ) %>%
  filter(!is.na(radius), radius > 0)
```

**Ex. 7.**

```{r}
nrow(long) # 27 635
```

**Ans.: 27,635 rows after filtering out NA and values ≤ 0.**

**Ex. 8.**

```{r}
long %>%
  tidyr::separate(quadrant, into = c("threshold", "direction"), sep = "_(?=[A-Z]{2}$)") %>%
  filter(threshold == "r50") %>%
  group_by(direction) %>%
  summarise(mean_radius = mean(radius, na.rm = TRUE)) %>%
  arrange(desc(mean_radius))
```

**Ans.: NE typically has the largest radius (the right-front quadrant is the most dangerous).**

**Ex. 9.**

```{r}
hur %>%
  group_by(year) %>%
  summarise(max_wind = max(wind_kts, na.rm = TRUE)) %>%
  arrange(desc(max_wind)) %>%
  slice(1)

# Which hurricane in that year?
hur %>%
  filter(year == 1988) %>%
  filter(wind_kts == max(wind_kts, na.rm = TRUE)) %>%
  select(storm_name, wind_kts, year)
```

**Ans.: Year 1988 (GILBERT, 160 kt).**

**Ex. 10.**

```{r}
hur %>%
  group_by(storm_id, storm_name) %>%
  summarise(mean_wind = mean(wind_kts, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(mean_wind)) %>%
  slice(1:3)
```

**Ex. 11.** → **the widths (in characters) of successive columns and their names.**

**Ex. 12.** → **column names a…z become values in the column `variable`.**

**Ex. 13.** → **missing measurement data** (-99 is the standard "missing" placeholder in this database).

**Ex. 14.**

```{r}
hur %>%
  mutate(kategoria = case_when(
    wind_kts >= 137 ~ "Cat5",
    wind_kts >= 96 ~ "Cat3",
    wind_kts >= 64 ~ "Cat1",
    wind_kts >= 34 ~ "TS",
    TRUE ~ "below_TS"
  )) %>%
  count(kategoria) %>%
  arrange(desc(n))
```

**Ans.: Most are TS (34–63 kt); Cat1–Cat5 make up ~25% of observations.**
